---
title: "allplayer"
format: html
editor: visual
---

## Introduction:

This code is all about player shooting performance between Washington wizards and Boston Celtics.

It displays all the interactive graphs which visualize and compare the shooting performance of Washington wizards and Boston Celtics players  

## Running Code

```{r message=TRUE, warning=TRUE}
#| code-fold: true

library(plotly)
library(readxl)
```

```{r}
#| code-fold: true

data <- read_excel("Data23-24/allplayer_23-24.xlsx")
head(data)
```

```{r echo=FALSE}
#| code-fold: true

wizards_players <- subset(data, Tm == "WAS")
celtics_players <- subset(data, Tm == "BOS")
```

```{r echo=FALSE}
#| code-fold: true

# Create a scatter plot
plot <- plot_ly()

# Add traces for each team
plot <- add_trace(plot, x = wizards_players$PTS, y = wizards_players$AST, mode = "markers", type = "scatter", name = "Washington Wizards")
plot <- add_trace(plot, x = celtics_players$PTS, y = celtics_players$AST, mode = "markers", type = "scatter", name = "Boston Celtics")

# Customize the layout
plot <- layout(plot, title = "Player Shooting Performance Comparison",
               xaxis = list(title = "PTS"),
               yaxis = list(title = "AST"))

# Show the plot
plot
```

```{r echo=FALSE}
#| code-fold: true

# Filter out rows with missing shooting statistics
wizards_players <- na.omit(wizards_players[, c("Player", "FG%", "3P%", "FT%")])
celtics_players <- na.omit(celtics_players[, c("Player", "FG%", "3P%", "FT%")])
# Create an interactive stacked bar chart
plot <- plot_ly() %>%
  add_trace(data = wizards_players, x = ~Player, y = ~`FG%`, name = 'FG% - Washington Wizards', type = 'bar') %>%
  add_trace(data = celtics_players, x = ~Player, y = ~`FG%`, name = 'FG% - Boston Celtics', type = 'bar') %>%
  add_trace(data = wizards_players, x = ~Player, y = ~`3P%`, name = '3P% - Washington Wizards', type = 'bar') %>%
  add_trace(data = celtics_players, x = ~Player, y = ~`3P%`, name = '3P% - Boston Celtics', type = 'bar') %>%
  add_trace(data = wizards_players, x = ~Player, y = ~`FT%`, name = 'FT% - Washington Wizards', type = 'bar') %>%
  add_trace(data = celtics_players, x = ~Player, y = ~`FT%`, name = 'FT% - Boston Celtics', type = 'bar')

# Customize layout
plot <- layout(plot, title = "Player Shooting Performance Comparison",
               xaxis = list(title = "Player"),
               yaxis = list(title = "Shooting Percentage"),
               barmode = 'stack')

# Show the plot
plot
```

```{r}
#| code-fold: true

# Define custom colors for each team
wizards_color <- "#1E90FF"  # Blue
celtics_color <- "#008000"  # Green

# Create a plotly line chart
plot_ly() %>%
  add_lines(data = wizards_players, x = ~Player, y = ~`3P%`, color = I(wizards_color), name = "Wizards") %>%
  add_lines(data = celtics_players, x = ~Player, y = ~`3P%`, color = I(celtics_color), name = "Celtics") %>%
  layout(title = "Three-Point Shooting Performance: Washington Wizards vs Boston Celtics",
         xaxis = list(title = "Player Name"),
         yaxis = list(title = "Three-Point Percentage"),
         legend = list(x = 1, y = 1, xanchor = "right", yanchor = "top", orientation = "h"))  # Set legend position to top right
```

```{r}
#| code-fold: true

# Define custom colors for each team
wizards_color <- "#FF5733"  # Red
celtics_color <- "#3366CC"  # Blue

# Create a plotly bar chart
plot_ly() %>%
  add_trace(data = wizards_players, x = ~Player, y = ~`3P%`, name = "Washington Wizards", type = 'bar', marker = list(color = wizards_color)) %>%
  add_trace(data = celtics_players, x = ~Player, y = ~`3P%`, name = "Boston Celtics", type = 'bar', marker = list(color = celtics_color)) %>%
  layout(title = "Three-Point Shooting Performance: Washington Wizards vs Boston Celtics",
         xaxis = list(title = "Player Name"),
         yaxis = list(title = "Three-Point Percentage"),
         barmode = 'group')

```

When we turn our gaze to the forwards, the contrast in shooting performance becomes even more pronounced. Kyle Kuzma and Jayson Tatum, (these two player all play a roles as leading representing their respective teams, have been dissected through Points Per Shot (PPS) statistics to illustrate this disparity.

```{r}
#| code-fold: true

library(tidyverse)
library(mgcv)
library(nbastatR)

#Double size to acquire shot data
Sys.setenv("VROOM_CONNECTION_SIZE" = 131072 * 2)

#Source data for court plotting
source("https://raw.githubusercontent.com/toddwschneider/ballr/master/plot_court.R")
source("https://raw.githubusercontent.com/toddwschneider/ballr/master/court_themes.R")

# Get all NBA teams
nba_teams <- nbastatR::nba_teams() %>% dplyr::filter(isNonNBATeam == 0,
                                                    idDivision > 0)

#Acquire shot data
season_shots_list <- 1:nrow(nba_teams) %>%
  purrr::map(function(x) nbastatR::teams_shots(team_ids = nba_teams$idTeam[x],
                                               seasons = 2024), .progress = T)

#Create shot locations + FG% + Points Per Shot
season_shots <- season_shots_list %>% bind_rows() %>%
  mutate(shot_attempt = ifelse(isShotAttempted == T,1,0),
         shot_made = ifelse(isShotMade == T,1,0),
         loc_x = -as.numeric(as.character(locationX))/10,
         loc_y = as.numeric(as.character(locationY))/10 + hoop_center_y,
         pps = case_when(typeShot == "2PT Field Goal" & shot_made == 1 ~ 2,
                         typeShot == "3PT Field Goal" & shot_made == 1 ~ 3,
                         typeShot == "2PT Field Goal" & shot_made == 0 ~ 0,
                         typeShot == "3PT Field Goal" & shot_made == 0 ~ 0,
                         TRUE ~ NA_real_))


season_shot_pct <- season_shots %>% group_by(zoneBasic) %>%
  summarise(makes = sum(shot_made),
            attempts = sum(shot_attempt),
            pps = sum(pps)) %>%
  ungroup() %>% mutate(lg_xfg_pct = round(makes/attempts,3),
                       lg_pps = round(pps/attempts,3)) %>%
  select(zoneBasic,lg_xfg_pct,lg_pps)

#Create GAM function for xFG% and PPS
xfg_aa_gam_fit <- function(d) {
  gam(as.numeric(xfg_pct_aa) ~
        s(loc_x, loc_y),
      data = d)
}

pps_aa_gam_fit <- function(d) {
  gam(as.numeric(pps_aa) ~
        s(loc_x, loc_y),
      data = d)
}

#Create prediction grid
grid_predict <- function(fit){
  grid <- expand.grid(loc_x = seq(-24,24,length=100),
                      loc_y = seq(0,32,length=100))
  grid$lp <- predict(fit,grid, type = "response")
  
  grid
}

#Heatmap
shot_heatmap_chart <- function(player, court_theme = court_themes$dark,
                               type = "PPS") {
  
  if (type == "PPS") {
    p_title <- paste(player,"PPS Above Avg Heatmap")
    plot_lims <- c(-1,1)
    plot_breaks <- c(-1,0,1)
  } else {
    p_title <- paste(player,"xFG% Above Avg Heatmap")
    plot_lims <- c(-0.3,0.3)
    plot_breaks <- c(-0.3,0,0.3)
  }
  
  df <- season_shots %>% dplyr::filter(namePlayer == player)
  
  shot_pct <- df %>% group_by(zoneBasic) %>%
    summarise(makes = sum(shot_made),
              attempts = sum(shot_attempt),
              pps = sum(pps)) %>%
    ungroup() %>% mutate(xfg_pct = round(makes/attempts,3),
                         pps = round(pps/attempts,3)) %>%
    select(zoneBasic,xfg_pct,pps)
  
  if (type == "PPS") {
    df <- df %>% select(-pps) %>%
      left_join(.,shot_pct,by="zoneBasic") %>%
      left_join(.,season_shot_pct,by="zoneBasic") %>%
      mutate(pps_aa = pps - lg_pps,
             xfg_pct_aa = xfg_pct - lg_xfg_pct) %>%
      pps_aa_gam_fit() %>%
      grid_predict()
  } else {
    df <- df %>% select(-pps) %>%
      left_join(.,shot_pct,by="zoneBasic") %>%
      left_join(.,season_shot_pct,by="zoneBasic") %>%
      mutate(pps_aa = pps - lg_pps,
             xfg_pct_aa = xfg_pct - lg_xfg_pct) %>%
      xfg_aa_gam_fit() %>%
      grid_predict()
  }
  
  # plot_court() +
  #   geom_tile(data=df,aes(x=loc_x,y=loc_y,
  #                         fill = lp)) +
  #   scale_fill_viridis_c(
  #     "%     ",
  #     limits = plot_lims,
  #     breaks = plot_breaks,
  #     option = "inferno",
  #     guide = guide_colorbar(barwidth = 15)
  #   ) +
  #   geom_path(data = court_points,
  #             aes(x = x, y = y, group = desc),
  #             color = court_theme$lines) +
  #   xlim(-24,24) +
  #   ylim(0,32) +
  #   coord_fixed() +
  #   theme(legend.text = element_text(size = rel(0.6))) +
  #   ggtitle(label = p_title,
  #           subtitle = "2023-24 Season")
  plot_court() +
  geom_tile(data=df, aes(x=loc_x, y=loc_y, fill = lp)) +
  scale_fill_gradient2(
    low = "#00008B", high = "#B22222", mid = "#FFFFFF",
    midpoint = 0,
    limits = plot_lims,
    breaks = plot_breaks,
    labels = c("-1", "0", "1"), # Add labels for each break
    guide = guide_colorbar(barwidth = 15)
  ) +
  geom_path(data = court_points,
            aes(x = x, y = y, group = desc),
            color = court_theme$lines) +
  xlim(-24, 24) +
  ylim(0, 32) +
  coord_fixed() +
  theme(legend.text = element_text(size = rel(0.6))) +
  ggtitle(label = p_title, subtitle = "2023-24 Season")
  
}

#Example
shot_heatmap_chart("Jordan Poole", type = "PPS")
shot_heatmap_chart("Jordan Poole", type = "xFG%")
shot_heatmap_chart("Donovan Mitchell", type = "PPS")
shot_heatmap_chart("Donovan Mitchell", type = "xFG%")
shot_heatmap_chart("Kyle Kuzma", type = "PPS")
shot_heatmap_chart("Kyle Kuzma", type = "xFG%")
shot_heatmap_chart("Jayson Tatum", type = "PPS")
shot_heatmap_chart("Jayson Tatum", type = "xFG%")
```

In conclusion , we make several plots about showing how bad washington wizzard is. First plot is compare the overall performance between washington wizzard and rest 29 team mean performace. We can say that wizzard are a low basketball efficiency team, basically they shoot a lot and miss a lot. Also they lose average 10 points per game show as the second plot. Then we compare the player performance. Basically every player's performance is below average. And then we doing same detail shooting perfermance of wizzards. And we also compare the wizzard star player with the other star players. 

## **Limitations of the Current Analysis**
### **Data Scope and Availability**

One significant limitation of our analysis lies in the data's scope and availability. We primarily focused on , which, while crucial, do not encompass all aspects of a player's impact on the game. Defensive contributions, playmaking abilities, and other intangible elements like leadership and game IQ are not captured in these shooting metrics. Additionally, our analysis could be constrained by the temporal range of the data, as we have concentrated solely on the 2023-24 season. This snapshot provides limited insight into long-term trends and player development over time.

### **Reliance on Publicly Available Metrics**

Our project heavily relies on publicly available metrics like Points Per Shot (PPS) and expected Field Goal Percentage (xFG%). These metrics, while informative, can sometimes oversimplify complex game situations. For instance, they might not adequately account for the quality of opposition defenses or the context of shots taken (e.g., end of shot clock scenarios).

## **Future Directions**

### **Expanding Data Dimensions**

To build on the current analysis, future studies could incorporate more comprehensive data sets that include defensive statistics, plus/minus scores, win shares, player efficiency ratings, and advanced metrics that consider shot context. This expansion would provide a more holistic view of a player's overall effectiveness and contribution to the team's success.

### **Longitudinal Studies**

Conducting longitudinal studies over multiple seasons could also provide deeper insights into player development and team dynamics. This approach would allow us to identify patterns and trends, potentially leading to more predictive analytics that can forecast future performance based on past data.

### **Integration of Machine Learning**

Leveraging machine learning techniques could refine our analysis further. For instance, predictive models could be developed to forecast shooting performances based on a variety of factors, including player fatigue, opposition strength, and even psychological factors such as player confidence and team morale.

### **Real-Time Data Utilization**

Lastly, integrating real-time game data could revolutionize our interactive visualizations, allowing fans and analysts alike to see live updates and perhaps even predictive insights during actual games. This would not only enhance the user experience but also provide immediate contextual analysis, adding a dynamic layer to our interactive tools.
